---
title       : CIS 4930 Fall 2015 Data Minining
subtitle    : Final Project Roadmap
author      : Dax Gerts, Christine Moore, Carl Amko, Denzel Mathew, Aaron Silcott
date        : December 16th, 2015
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Read-And-Delete

# Project Tasks

* Dataset creation.
* Genre prediction (classification)
* The usual casts (association rule mining)
* Finding similar movies (clustering)

--- .class #id 

## Dataset Creation (1)

Several trajectories were taken in efforts to most effectively create a working dataset from the IMDB database.

1. Pure R Approach
  * R/RStudio
2. Database Approach
  * Wget
  * Python (IMDBpy)
  * SQL (imddbpy2sql.py)
  * CSV (sqldump)

--- .class #id

## Dataset Creation (2)

In efforts to make loading numerous large datasets in R more efficient, several helper functions were written.

### Download procedure in R

```{r, eval=TRUE,echo=FALSE,cache=TRUE}
ready <- FALSE
loadPackages1 <- function() {
  if( require(R.utils) == FALSE) { install.packages("R.utils") }
  ready <- TRUE
}
```
```{r, eval = TRUE}
temp <- tempfile() #Prepare space for downloading files
download.file("ftp://ftp.fu-berlin.de/pub/misc/movies/database/movies.list.gz",temp)
rawData <- readLines(temp)
unlink(temp) #Destroy temp file
rawData
```

--- .class #id

## Dataset Creation (3)

### Single-function downloader

```{r}
downloadRawDataset <- function(choice) {
  temp <- tempfile() #Prepare space for downloading files
  
  if(choice == "actors") { download.file("ftp://ftp.fu-berlin.de/pub/misc/movies/database/actors.list.gz",temp) }
  if(choice == "actresses") { download.file("ftp://ftp.fu-berlin.de/pub/misc/movies/database/actresses.list.gz",temp) }
  if(choice == "complete-cast") { download.file("ftp://ftp.fu-berlin.de/pub/misc/movies/database/complete-cast.list.gz",temp) }
  if(choice == "composers") { download.file("ftp://ftp.fu-berlin.de/pub/misc/movies/database/directors.list.gz",temp) }
  if(choice == "directors") { download.file("ftp://ftp.fu-berlin.de/pub/misc/movies/database/directors.list.gz",temp) }
  if(choice == "genres") { download.file("ftp://ftp.fu-berlin.de/pub/misc/movies/database/genres.list.gz",temp) }
  if(choice == "keywords") { download.file("ftp://ftp.fu-berlin.de/pub/misc/movies/database/keywords.list.gz",temp)}
#...etc
  
  rawData <- readLines(temp)
  unlink(temp)
  rawData
}
```

--- .class #id

## Dataset Creation (4)

### Download + explore values

```{r}
while(ready == FALSE) { ready <- loadPackages1() }
exploreData <- function(choice,length = 50) {
  dt <- downloadRawDataset(choice)
  print(head(dt,length))
  cat("Press [enter] to continue")
  line <- readline()
  print(tail(dt,length))
  dt
}
```

--- .class #id

## Genre Prediction (1)


--- .class #id 

## The *Usual* Casts (1)


--- .class #id 

## Finding Similar Movies (1)


--- .class #id 

## Conclusions


--- .class #id 

## Further Endeavors




